### 2017-08-25 paper/CoverLetter.tex

Studying multivariate data as an alternative to univariate data has become increasingly popular in the last decades both in ecology and evolution.
This includes, for example, studying form and function (e.g. Diaz et al., 2016, \textit{Nature}) or morphological diversity (i.e. disparity - e.g. Brusatte et al. 2012, \textit{Science}).
These methods have in common to use a multidimensional space (through ordination or other techniques) representing the combination of all possible observed traits and then measuring some aspect of the space (e.g. its hyper-volume) for different groups and compare them between each other or to a null hypothesis.

Hence there exist a multitude of ways to define a multidimensional space, a metric measuring a specific aspect thereof and a test for many hypothesis.
In theory this vast tool-kit should allow evolutionary biologists and ecologists to do many things.
There are numerous implementation of such analysis pipeline in numerous R packages (e.g. \texttt{Claddis}, \texttt{geomorph}, \texttt{ade4}, \texttt{vegan}, etc.) or even in other languages (e.g. \texttt{GINGKO} in javascript).
However, each of these implementations are particular to the software/package author's definition of the multidimensional space, metric and test.
This leads researchers to often use these tool-kits by default hampering them to use the best definitions of the space/metric/test fitted to their particular question.
At worst this leads in turn to wrong statistic claims, and, at best to time consuming code retro-engineering.

In this application manuscript entitled ``dispRity: an R package for measuring disparity'', I present a complete and flexible R package allowing to run multidimensional analysis where each parameter (the space, the metric and the test) can be easily defined by the user.
In fact, this package provides a tidy and modular environment where users can use their own multidimensional space defined as a matrix and apply to it any metric and any test (provided as functions).
This allows for highly modular multidimensional analysis and the tidy interface allows easily reproducible analysis.

>>>>

Studying multivariate data as an alternative to univariate data has become increasingly popular in the last decades in both ecology and evolution.
This includes, for example, studying form and function (e.g. Diaz et al., 2016, \textit{Nature}) or morphological diversity (i.e. disparity - e.g. Brusatte et al. 2012, \textit{Science}).
These methods all use a multidimensional space (through ordination or other techniques) representing the combination of all possible observed traits and then measuring some aspect of the space (e.g. its hypervolume) for different groups, and comparing these to each other or to a null hypothesis.

There are a multitude of ways of carrying out each of these steps, and numerous software implementations in several (e.g. in R: \texttt{Claddis}, \texttt{geomorph}, \texttt{ade4}, \texttt{vegan}, in javascript: \texttt{GINGKO}).
However, each of these implementations is particular to the package author's definition of the multidimensional space, metric and test.
This leads researchers to use these toolkits by default, thus preventing them from using the best definitions of the space/metric/test for their particular question or dataset.
At worst this leads to incorrect statistic claims, and, at best to time consuming code retro-engineering.

In this application manuscript entitled ``dispRity: an R package for measuring disparity'', I present a complete and flexible R package for running multidimensional analysis where each parameter can be easily defined by the user (the space, the metric and the test).
This package provides a tidy and modular environment where users can use their own multidimensional space defined as a matrix and apply to it any metric and any test (provided as functions).
This allows for highly modular multidimensional analysis and the tidy interface allows easily reproducible analysis.
All functions have documentation and unit tests. 
Finally, this package comes with an in depth Gitbook manual covering most functionalities that will be updated regularly following user requests.

% NC: Add maybe somewhere how many people are already using the package? The package is likely to be well cited as it is already being used in at least XX projects that I am aware of.
